Predicting Political Support in the United States

### 2024-03-17 output/paper.qmd
fix figure crossref and justification formatting, fix grammar mistakes

The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump [cite2020Election]. Using the 2022 Cooperative Election Study (CES) survey data, we aim to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors. We will be using a representative sample of 60000 Americans to discern patterns in these voters' responses and how their personal backgrounds may have influenced their voting decision. We will be examining the impact of various factors such as: gender, race, educational background, religon, and personal stances on key political issues such as gun control and student loans. Our main focus is to compare the correlations in the 2022 CES survey data to the 2020 CES survey data, and to see if our estimand, the effect of a voter's cultural background and political stances on their voting decision, has influenced their voting decision in the 2020 election. Using these understandings, we hope to highlight the potential implications these correlations may have for this upcoming 2024 election.

@@

The variables we used in our analysis were: - `presvote20post`, renamed to `voted_for` - This data was represented in the form of a numerical variable corresponding to the presidential candidate that voted for in the 2020 eleciton. This was cleaned to limit the options to either "Biden" or "Trump", the two leading candidates in the 2020 and 2024 election.

@@

    Incorporating the voter's `religion`, was a needed factor since it is important to understand that religious beliefs and differences can have a strong influence one's personal beliefs and therefore impact their personal stances on topics such as gun ownership. To put a contrasting example out there, 30% of Protestant/Christian respondents support the Democrats as opposed to 72% by Atheists. This could be attributed to certain parties aligning more to certain religious beliefs. To see a bigger insight into this, take a look at @fig-racevsreligionlarge and @fig-racevsreligion2 (figures separated according to respondent count), and notice examples like, 'White Protestants' appear significantly more supportive towards the Republicans as opposed to 'White Catholics' and 'Black Protestants' who appear to be significantly more supportive towards the Democrats.

@@

# Model

@@

## Political Importances Model

@@

  @tbl-modelsummarycultural first suggests a that males were less likely to vote for Biden, and more likely to vote for Trump. Looking at @fig-agevsgender-1 this is not an obvious correlation to make with the votes for Biden and Trump being close for each age group, with Trump only winning more votes with the male and above 65 demographic. However, looking at the 90% confidence intervals in @fig-conf_cultural for the coefficients of that model, specifically that for being male, we can see that the model is 90% confident that this relationship is around strictly negative against voting for Biden. We also see the disagreement within the 45+ community that @fig-agevsgender-1 shows, and the agreement within the younger community (18 - 45), with @tbl-modelsummarycultural showing a large correlation between being 30-44 to supporting Biden, with being aged 45-65 or 65+ having very little correlation to support Biden, supported by the 90% confidence interval shown in @fig-conf_cultural for those coefficients spanning both negative and positive, suggesting an uncertain preference. Although the model included being: Female, Asian, 18-29, and Agnostic, in the intercept, for @fig-agevsgender-1, @fig-racevsreligion2, and @fig-racevsreligionlarge and @fig-racevsreligion2 respectively, it reasonable to say that each graph suggests that being Female, Asian, 18-29, and Agnostic has a correlation with voting for Biden, each demographic having a significant amount more votes for Biden than Trump.

@@

weakenesses of the analysis::: TODO also talk about small demographic sizes

>>>>

The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump \[cite2020Election\]. Using the 2022 Cooperative Election Study (CES) survey data, we aim to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors. We will be using a representative sample of 60000 Americans to discern patterns in these voters' responses and how their personal backgrounds may have influenced their voting decision. We will be examining the impact of various factors such as: gender, race, educational background, religion, and personal stances on key political issues such as gun control and student loans. Our main focus is to compare the correlations in the 2022 CES survey data to the 2020 CES survey data, and to see if our estimand, the effect of a voter's cultural background and political stances on their voting decision, has influenced their voting decision in the 2020 election. Using these understandings, we hope to highlight the potential implications these correlations may have for this upcoming 2024 election.

@@

The variables we used in our analysis were: - `presvote20post`, renamed to `voted_for` - This data was represented in the form of a numerical variable corresponding to the presidential candidate that voted for in the 2020 election. This was cleaned to limit the options to either "Biden" or "Trump", the two leading candidates in the 2020 and 2024 election.

@@

Incorporating the voter's `religion`, was a needed factor since it is important to understand that religious beliefs and differences can have a strong influence one's personal beliefs and therefore impact their personal stances on topics such as gun ownership. To put a contrasting example out there, 30% of Protestant/Christian respondents support the Democrats as opposed to 72% by Atheists. This could be attributed to certain parties aligning more to certain religious beliefs. To see a bigger insight into this, take a look at @fig-racevsreligionlarge-1, @fig-racevsreligionlarge-2 and @fig-racevsreligion2 (figures separated according to respondent count), and notice examples like, 'White Protestants' appear significantly more supportive towards the Republicans as opposed to 'White Catholics' and 'Black Protestants' who appear to be significantly more supportive towards the Democrats.

@@

# Model {#model}

@@

## Political Importance Model

@@

@tbl-modelsummarycultural first suggests a that males were less likely to vote for Biden, and more likely to vote for Trump. Looking at @fig-agevsgender-1 this is not an obvious correlation to make with the votes for Biden and Trump being close for each age group, with Trump only winning more votes with the male and above 65 demographic. However, looking at the 90% confidence intervals in @fig-conf_cultural for the coefficients of that model, specifically that for being male, we can see that the model is 90% confident that this relationship is around strictly negative against voting for Biden. We also see the disagreement within the 45+ community that @fig-agevsgender-1 shows, and the agreement within the younger community (18 - 45), with @tbl-modelsummarycultural showing a large correlation between being 30-44 to supporting Biden, with being aged 45-65 or 65+ having very little correlation to support Biden, supported by the 90% confidence interval shown in @fig-conf_cultural for those coefficients spanning both negative and positive, suggesting an uncertain preference. Although the model included being: Female, Asian, 18-29, and Agnostic, in the intercept, for @fig-agevsgender-1, @fig-racevsreligion2, and @fig-racevsreligionlarge-1, @fig-racevsreligionlarge-2 and @fig-racevsreligion2 respectively, it reasonable to say that each graph suggests that being Female, Asian, 18-29, and Agnostic has a correlation with voting for Biden, each demographic having a significant amount more votes for Biden than Trump.

@@

weaknesses of the analysis::: TODO also talk about small demographic sizes
